---
title: "Cabbages"
author: "Anjolie Barrios"
date: "Feb 8, 2021"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(MASS)
library(dplyr)
veggy <- cabbages %>%
  mutate(vitC_density = VitC / HeadWt) 
  c39 <- subset(veggy,Cult=="c39") 
  c52 <- subset(veggy,Cult=="c52")
```

# Introduction
<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->
 The cabbages data from the MASS package is loaded into a dataframe called veggy. Complete this report by answering the following two questions  
This data comes from tests of cabbages with c39 and c52 cultivation methods; using this data, we'll find out which, if any, cultivation method is better.  
  
### Is there a difference in average VitC between c39 and c52 cultivation methods?  
  
$H_0$: There is no difference in average VitC between c39 and c52 cultivation methods.  
  
$H_A$: There is a difference in average VitC between c39 and c52 cultivation methods.  
  
# Methods  
<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
Since cultivation is categorical and VitC is numerical, a Num~Cat variable analysis will be used.  
A boxplot will be the graphical method. A t- test and summary will be the numerical methods.  

# Results  

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results  

### Graphical Descriptive Results  
```{r}
plot(VitC~Cult,data=veggy) %>%
  abline(h = 51.5, lwd = 2, col = "red") %>%
  abline(h = 64.4, lwd = 2, col = "blue") %>%
  title(main = "Vitamin C content vs Cultivation method") %>%
  title(xlab="Cultivation Method", mgp=c(2,1,0)) %>%
  title(ylab="Vitamin C levels", mgp=c(2,1,0))
```
  
The red line is the average VitC content of c39 cultivation (51.5), and the blue line is that of c52 cultivation (64.4). Units are unknown.  
  
```{r}
plot(vitC_density~Cult,data=veggy) %>%
  abline(h = 19.66936, lwd = 2, col = "red") %>%
  abline(h = 33.02514, lwd = 2, col = "blue") %>%
  title(main = "Vitamin C Density") %>%
  title(xlab="Cultivation Method", mgp=c(2,1,0)) %>%
  title(ylab="Vitamin C density", mgp=c(2,1,0))
```
  
Density was calculated by dividing the Vitamin C content by weight. Units are unknown.  
The red line is the average VitC density of c39 cultivation (19.66936), and the blue line is that of c52 cultivation (33.02514). A greater disparity in VitC content (both between and within each category of cultivation) is found when density is taken into account, but c52 cultivation beats c39 either way.  
```{r}
model.1 <- lm(VitC ~ HeadWt, data = veggy)
data1 <- ggplot(veggy,aes(y=VitC,x=HeadWt))
data1 + geom_point() + geom_smooth(method = lm)
```
  
As seen in this graph, head weight and VitC content are negatively correlated. However, it is unknown at this point whether more VitC content is a direct result of the cultivation method or whether, regardless of cultivation, VitC is less common in heavier cabbages.  
<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Numerical Descriptive Results  
<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
favstats(VitC~Cult,data=veggy)
```
  
64.4 > 51.5, which proves c52 yields more VitC per cabbage head on average. 
  
```{r}
favstats(vitC_density~Cult,data=veggy)
```
  
33.0251 > 19.6694, which proves c52 yields more VitC per weight on average. The greater disparity in averages can be explained thus:  
```{r}
favstats(HeadWt~Cult,data=veggy)
```
  
On average, c39 cabbages are heavier, so their VitC content is spread over a greater weight or mass.  

## Inferential Results  
<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
The alternate hypothesis is that there is a difference in average VitC between c39 and c52 cultivation methods; the null hypothesis is that there is NO difference in average VitC between c39 and c52 cultivation methods.  
The following parametric t- test was chosen because the data fulfilled two assumptions: the VitC variable has similar variances (50.7414 and 71.4897) in both groups, and they're interpretted in a ratio scale (with an absolute zero and able to be used to make meaningful ratios). The data isn't normally distributed or big enough to represent an entire population, though.  
The variable "vitC_density" was not used in tests because its variances (~70 and ~216 respectively) are too far apart.  
```{r}
t.test(VitC~Cult,data=veggy)
```
  
If the null hypothesis was true, there would be a 0.00000003405 probability of these results happening. That's less than 0.05, so the null hypothesis is probably false. 

##HeadWt vs VitC content

```{r}
model.c39 <- lm(VitC ~ HeadWt, data = c39)
summary.lm(model.c39)
```

```{r}
model.c52 <- lm(VitC ~ HeadWt, data = c52)
summary.lm(model.c52)
```
  
The slope for the line of regression is -4.526 for c39; it's -6.647 for c52, so HeadWt and VitC are more negatively correlated in c52 cabbages.  

# Discussion and Conclusion  
While the lack of units limits the usefulness of this data, it can be assumed that (in cabbage), on average, c52 cultivation yields more Vitamin C than c39 cultivation.  
However, it is unknown whether more VitC content is a direct result of the cultivation method or whether, regardless of cultivation, growing bigger cabbages results in less VitC content. If growing bigger cabbages is an inherent part of c39, then it's unknown whether the effect is direct or indirect (as a result of the cabbages being bigger).  If not, results are still inconclusive on the cause of c52 cabbages having more VitC. 
<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
